Environment modeling in \tool

### 2022-05-02 paper.tex
pass abstract and intro

\begin{abstract}
  Symbolic execution is a popular program analysis technique for software security and verification. It aims to explore multiple feasible execution of a program in parallel while gathering the corresponding path constraints. As real-world programs often require interaction with the underlying operating system, modeling the environment plays a crucial role in symbolic execution. Coreutils is a collection of the basic file, shell manipulation commands of the GNU operating system. It is used by most symbolic execution framework as a stress test benchmark for intense environment interactions. \par
  \tool is a symbolic execution compiler for LLVM IR that realizes efficient execution by compilation. In this paper, we first present two different approaches of environment modeling for \tool. Then we demonstrate various obstacles we overcome towards testing Coreutils in \tool. Finally we compare our framework with other symbolic engine based on benchmarks of Coreutils.

@@

Symbolic execution is widely used in software verification, bug-finding and automatic test generation. Although traditional unit testing is sufficient for most scenarios, it is still a under-approximation due to its inherent nature and requires increasing maintenance as program develops. First introduced in the 1970s\cite{boyer1975select, howden1977symbolic, king1975new, king1976symbolic}, symbolic execution can simultaneously explore different execution paths by making certain input symbolic and accompanying each path with a state. A state in symbolic execution engine is composed of 2 parts, a set of boolean formula to describe the path condition and a symbolic store that map variables to both symbolic and concrete values. Execution within a basic block will update the symbolic store the same as   the corresponding native execution. The executing path will fork when a branch with symbolic condition is encountered, generate two paths represent the true and false branch, check their feasibility with solver and update path condition accordingly. With the development of SMT\cite{barrett2021satisfiability} solvers like Z3\cite{moura2008z3}, symbolic execution is becoming increasingly practical and efficient.\par

@@

Traditional symbolic execution engine like KLEE\cite{cadar2008klee} is interpreter-based. The interpreter will traverse the compiled intermediate representation (IR) of the program, mapping each instruction into its symbolic semantic and update the path condition and symbolic store within their customized state representation according to the strategy we described above. Although interpreter-based symbolic execution can detect all potential bugs in theory, problems like path-explosion \cite{baldoni2018survey} prevent it from scaling up to large programs. The interpretation overhead of translating instruction into its symbolic semantic will also cause a significant performance slow-down.\par
Another approach to implement symbolic execution is instrumentation\cite{10.1145/1180405.1180445, CREST}. Instrumentation does not require a internal state-representation and avoid the interpretation overhead. Instructions on concrete values is executed natively while execution on symbolic values is achieved through instrumented code. Instrumentation naturally works well with concolic execution \cite{godefroid2005dart, sen2005cute} where only a single path is reasoned about at a time. While instrumentation achieves better performance compared to interpretation, its lack of state-representation prevents it from adopting certain optimization like state-merging and path selection strategies. \par
\tool is the first symbolic execution compiler framework that adapts compiler and code generation technique for symbolic execution. It eliminates the overhead of Interpretation while give user fine-grained control over the state representation, and achieves flexible and efficient symbolic execution. Various optimization and heuristic can be easily adopted in \tool by inspecting and manipulation over the internal state and control representation. The core idea of \tool is to specialize a symbolic interpreter with the input program using Partial evaluation\cite{damian1999partial}. According to the first Futamura projection\cite{Futamura1971, Futamura1999}, the specialized program is equivalent to running the interpreter with the input code. \tool achieves partial evaluation for the symbolic interpreter using staging\cite{DBLP:conf/pepm/TahaS97} and avoid the need to manually developing a complex specializer. With the symbolic interpreter annotated with stages, the program can be easily specialized using meta-programming, and we can get a symbolic execution compiler for free. Based on LMS\cite{rompf2010lightweight} - a library-based staging compiler that distinguishes stages solely on Type, \tool specializes a staged symbolic interpreter into a compiler and eliminates the interpretation overhead.

@@

As most programs are not self-contained, they need to interact with the underlying environment (operating system, network). Different flavors of symbolic execution have different treatment for external calls. \par
Instrumentation based concolic execution can directly invoke the native external call as each path is executed concretely. This approach is simple and straightforward but sacrifice the completeness of exploration. An alternative approach is to create an abstract model of environment accompanying different execution paths. As modeling external call at the library level is expensive, most framework chooses to model the environment at the system call level.
 KLEE\cite{cadar2008klee} develops a core symbolic file system where different paths can invoke system call in isolation. \par
The approaches above only investigates which part of the environment should be modeled but neglect the question about how to model the environment. As we mentioned above, \tool is a symbolic execution compiler that translate the programs in Source Language S to programs in Target language T. Intuitively, we can choose to model the environment in both the Source Language level (as part of the user program) and the Target Language level (as runtime library). Like KLEE, we choose to model the environment in the system call level with a symbolic file system. Using the flexibility of our framework, we can choose between directly poring KLEE' file system or developing a internal file system in the backend.\par
Coreutils \cite{coreutilsweb} is a collection of commonly used GNU command that manipulate file system, shell and text. It interacts intensively with the operating system and is used by most symbolic execution framework as benchmarks. We also use Coreutils as a stress test and benchmark for our environment modeling implementation.

@@

In section \ref{coreutilstests}, we will discuss various obstacles and performance bottlenecks we observed in  adapting Coreutils tests into our engine and our corresponding solutions.\par

@@

\section{Different levels of Environment Modeling}\label{twofsimpl}
In the context of symbolic execution compiler, we can choose between modeling the environment in the source language level or in the target language level. \tool translates LLVM-IR (compiled from C programs) into backend C++ programs. As we mentioned before, we choose to model the environment in the file system level like KLEE, so we can either directly porting KLEE's file system (written in C) or develops our own internal file system as backend C++ runtime library. We will discuss this two implementations in detail and compare their advantage and disadvantage.
\paragraph*{Porting KLEE's file system}
KLEE develops a POSIX-like symbolic file system on top of the native file system in order to alleviate side effects caused by external calls and prevent interference across execution paths. The symbolic file system is written purely in C and can be linked with the user program to be analyzed. This file system can be easily modified and extended without mangling
 with the internal details. In fact, by implementing the entire symbolic file system in the C level, it can be ported to other well-formed symbolic engine with no modification. We can directly port KLEE's file system into our model with little effort.\par
 KLEE's file system is a hybrid approach, it contains both symbolic files and concrete files. It maintains symbolic meta data like offset and open flags for both symbolic and concrete files. For symbolic files, it will allocated a char array buffer upon creation to store its symbolic content and a symbolic stat struct to store its symbolic status. For concrete file, it will records its actual file descriptor upon open for future references.\par
\begin{figure}[t]
\centering
\begin{subfigure}{0.5\textwidth}
\begin{lstlisting}[style=small,language=C,numbers=left,stepnumber=1,xleftmargin=2em,numberstyle=\ttfamily]
ssize_t write(int fd, const void *buf, size_t count) {
  exe_file_t *f;
  f = get_file(fd);

>>>>

\begin{abstract}
  Symbolic execution is a popular program analysis technique for software security and verification. It aims to explore multiple feasible executions of a program in parallel while gathering the corresponding path constraints. As real-world programs often require interaction with the underlying operating system, modeling the environment plays a crucial role in symbolic execution. Coreutils is a collection of the basic file, shell manipulation commands of the GNU operating system. Most symbolic execution frameworks use it as a stress test benchmark for intense environment interactions. \par
  \tool is a symbolic execution compiler for LLVM IR that realizes efficient execution by compilation. In this paper, we first present two different environment modeling approaches for \tool. Then we demonstrate various obstacles we overcome towards testing Coreutils in \tool. Finally, we compare these two file system implementations of \tool with another symbolic execution engine KLEE based on benchmarks of Coreutils.

@@

Symbolic execution is widely used in software verification, bug-finding, and automatic test generation. Although traditional unit testing is sufficient for most scenarios, it is still under-approximation due to its inherent nature and requires increasing maintenance as the program develops. First introduced in the 1970s\cite{boyer1975select, howden1977symbolic, king1975new, king1976symbolic}, symbolic execution can simultaneously explore different execution paths by making some inputs symbolic and accompanying each path with a state. A state in a symbolic execution engine is composed of 2 parts, a set of boolean formula to describe the path condition and a symbolic store that map variables to both symbolic and concrete values. Execution within a basic block will update the symbolic store the same as the corresponding native execution. The executing path will fork when a branch with a symbolic condition is encountered, generating two paths representing the true and false branch. The feasibility of both branches will be checked using a solver, and the path conditions will be updated accordingly. With the development of SMT\cite{barrett2021satisfiability} solvers like Z3\cite{moura2008z3}, symbolic execution is becoming increasingly practical and efficient.\par

@@

Traditional symbolic execution engine like KLEE\cite{cadar2008klee} is interpreter-based. The interpreter will traverse the program's intermediate representation (IR), mapping each instruction into its symbolic semantic and updating the path condition and symbolic store within their customized state representation according to the strategy we described above. Although interpreter-based symbolic execution can detect all potential bugs in theory, problems like path-explosion \cite{baldoni2018survey} prevent it from scaling up to large programs. The interpretation overhead of translating instruction into its symbolic semantic will also cause a significant performance slow-down.\par
Another approach to implementing symbolic execution is instrumentation\cite{10.1145/1180405.1180445, CREST}. Instrumentation does not require an internal state representation and avoids the interpretation overhead. Instructions on concrete values are executed natively, while execution on symbolic values is achieved through instrumented code. Instrumentation naturally works well with concolic execution \cite{godefroid2005dart, sen2005cute} where only a single path is reasoned about at a time. While instrumentation achieves better performance than interpretation, its lack of state representation prevents it from adopting optimizations like state-merging and some path selection strategies. \par
\tool is the first symbolic execution compiler framework that adopts compiler and code generation techniques for symbolic execution. It eliminates the overhead of Interpretation while giving the user fine-grained control over the state representation and achieves flexible and efficient symbolic execution. Various optimization and heuristic can be easily adopted in \tool by inspecting and manipulating the internal state and control representation. The core idea of \tool is to specialize a symbolic interpreter with the input program using Partial evaluation\cite{damian1999partial}. According to the first Futamura projection\cite{Futamura1971, Futamura1999}, the specialized program is equivalent to running the interpreter with the input code. \tool achieves partial evaluation for the symbolic interpreter using staging\cite{DBLP:conf/pepm/TahaS97} and avoids the need to develop a complex specializer manually. Based on LMS\cite{rompf2010lightweight} - a library-based staging compiler that distinguishes stages solely on Type, \tool specializes a staged symbolic interpreter into a compiler and eliminates the interpretation overhead. With the symbolic interpreter annotated with stages, we can easily specialize the program using meta-programming and get a symbolic execution compiler for free.

@@

Most programs are not self-contained. They need to interact with the underlying environment (operating system, network). Different flavors of symbolic execution have different treatments for external calls. \par
Instrumentation-based concolic execution can directly invoke the native external call as each path is executed concretely. This approach is simple but sacrifices the completeness of exploration. An alternative approach is to create an abstract model of the environment accompanying different execution paths. As modeling external call at the library level is expensive, most framework chooses to model the environment at the system call level.
 KLEE\cite{cadar2008klee} develops a core symbolic file system where different paths can invoke system calls in isolation. \par
Approaches above only investigate which part of the environment should be modeled but neglect the question of how to model the environment. \tool is a symbolic execution compiler that translates the programs from Source Language S to Target language T. Intuitively, we can choose to model the environment in the Source Language (as part of the user program) or in the Target Language (as runtime library). Like KLEE, we choose to model the environment in the system call level with a symbolic file system. Using the flexibility of our framework, we can choose between directly poring KLEE' file system or developing an internal file system in the backend.\par
Coreutils \cite{coreutilsweb} is a collection of commonly used GNU commands that manipulate the file system, shell, and text. It interacts intensively with the operating system and is used by most symbolic execution frameworks as benchmarks. We also use Coreutils as a stress test and benchmark for our environment modeling implementation.

@@

In section \ref{coreutilstests}, we will discuss various obstacles and performance bottlenecks we observed in adopting Coreutils tests into our engine and our corresponding solutions.\par

@@

\section{Different levels of Environment Modeling}\label{twofsimpl}
In the context of symbolic execution compiler, we can choose between modeling the environment in the source language level or in the target language level. \tool translates LLVM-IR (compiled from C programs) into backend C++ programs. As we mentioned before, we choose to model the environment in the file system level like KLEE, so we can either directly porting KLEE's file system (written in C) or develops our own internal file system as backend C++ runtime library. We will discuss this two implementations in detail and compare their advantage and disadvantage.
\paragraph*{Porting KLEE's file system}
KLEE develops a POSIX-like symbolic file system on top of the native file system in order to alleviate side effects caused by external calls and prevent interference across execution paths. The symbolic file system is written purely in C and can be linked with the user program to be analyzed. This file system can be easily modified and extended without mangling
 with the internal details. In fact, by implementing the entire symbolic file system in the C level, it can be ported to other well-formed symbolic engine with no modification. We can directly port KLEE's file system into our model with little effort.\par
 KLEE's file system is a hybrid approach, it contains both symbolic files and concrete files. It maintains symbolic meta data like offset and open flags for both symbolic and concrete files. For symbolic files, it will allocated a char array buffer upon creation to store its symbolic content and a symbolic stat struct to store its symbolic status. For concrete file, it will records its actual file descriptor upon open for future references.\par
\begin{figure}[t]
\centering
\begin{subfigure}{0.5\textwidth}
\begin{lstlisting}[style=small,language=C,numbers=left,stepnumber=1,xleftmargin=2em,numberstyle=\ttfamily]
ssize_t write(int fd, const void *buf, size_t count) {
  exe_file_t *f;
  f = get_file(fd);
